---
title: "Final Project"
output: pdf_document
---

```{r}
library(dplyr)
library(zoo) 
library(lubridate) 
library(mgcv)
library(TSA) 
library(dynlm)
library(tseries)
library(xts)
library(astsa)
library(rugarch)
```
```{r}
atl = read.csv("Atlanta_data.csv", head = TRUE)
atl
```


```{r}
atl <- read.csv("Atlanta_data.csv", head = TRUE)
atl$time<-as.Date(atl$time,format='%Y-%m-%d')
#temp <- ts(atl$tavg, start=c(2002,4,1), freq=365)
temp = xts(atl$tavg, atl$time)
temp = apply.weekly(temp, mean, na.rm=TRUE)
temp <- ts(temp, start=c(2002,4,1), freq=52)
# plot time series
ts.plot(temp,ylab="daily temprature", main="Atlanta Weekly Temprature Time Series")
acf(temp, main="ACF:Atlanta Weekly Temprature", na.action = na.pass, lag.max=52*2)
```
The plot of daily temperature value clearly shows cyclic patterns indicating seasonality in the data which is expected as the temperature is known to be dependent on the time of the year. In terms of trend there is no clear upward or downward in the original time series, however the ACF plot shows that there might be a trend in the data as well. Hence it would be worth fitting trend and seasonality components to the data and perform EDA on the residuals.



```{r}
time.pts = c(1:length(temp))
time.pts = c(time.pts - min(time.pts))/max(time.pts)

## splines regression to fit daily data 
temp.gam = gam(temp~s(time.pts))
temp.gam.fit = ts(fitted(temp.gam), start=c(2002,4,1), freq=52)

### overlap daily time series
plot(temp,ylab="daily temprature", main="Atlanta Daily Temprature Time Series & Trend") 
lines(temp.gam.fit,col="blue") 

temp.season = lm(temp~season(temp)-1)
temp.season.fit = ts(fitted(temp.season), start=c(2002,4,1), freq=52)

### overlap daily time series
plot(temp,ylab="daily temprature", main="Atlanta Daily Temprature Time Series & Seasonality") 
lines(temp.season.fit,col="red") 

## splines+ANOVA seasonality
temp.lm.gam = gam(temp~s(time.pts)+season(temp)-1)
temp.lm.gam.fit = ts(fitted(temp.lm.gam),start=c(2002,4,1), freq=52)

### overlap daily time series
plot(temp,ylab="daily temprature", main="Atlanta Daily Temprature Time Series & Trend+Seasonality") 
lines(temp.lm.gam.fit,col="green") 

# plot the residuals of models
resid <- residuals(temp.lm.gam)
resid.fit <- ts(resid,start=c(2002,4,1), freq=52) 
plot(resid.fit,ylab="daily temprature residuals", main="Atlanta Daily Temprature Residuals") 
acf(resid.fit, main="ACF:Atlanta Daily Temprature Residuals", na.action = na.pass, lag.max=52*2)
pacf(resid.fit, main="PACF:Atlanta Daily Temprature Residuals", na.action = na.pass, lag.max=52*2)
adf.test(resid)
adf.test(resid^2)
```



```{r}
n = length(temp)
n_forward=28
nfit = n-n_forward
temp.train = temp[1:nfit]
temp.test = temp[-(1:nfit)]

n = length(temp.train)

test_modelA <- function(p,d,q){
    mod = arima(temp.train, order=c(p,d,q), method="ML")
    current.aic <- AIC(mod)
    current.aic<-current.aic-2*(p+q+1)+2*(p+q+1)*n/(n-p-q-2)
    df = data.frame(p,d,q,current.aic)
    names(df) <- c("p","d","q","AIC")
    #print(paste(p,d,q,current.aic,sep=" "))
    return(df)
}

orders <- data.frame(Inf,Inf,Inf,Inf)
names(orders) <- c("p","d","q","AIC")
for (p in 0:10){
    for(d in 0:2) {
        for (q in 0:10) {
            possibleError <- tryCatch(
                orders<-rbind(orders,test_modelA(p,d,q)),
                error=function(e) {e}
            )
            if(inherits(possibleError, "error")) 
              next
        }
    }
}

orders <- orders[order(-orders$AIC),]
tail(orders)

AIC_v = orders$AIC 
AIC_v = sort(AIC_v, decreasing = TRUE)
AIC_v = AIC_v[-1]
plot(AIC_v, xlim = c(0,50),ylab = "AIC Values")
```



```{r}
Selected_Order = AIC_v[12]
orders[orders$AIC==Selected_Order,] 

daily_model <- arima(temp.train, order=c(4,0,3))
daily_model

select.resids <- residuals(daily_model)

acf(select.resids,main="ARIMA Residuals - Weekly Data",na.action = na.pass,lag.max=52*2)
acf(select.resids^2,main="ARIMA Squared Residuals - Weekly Data",na.action = na.pass,lag.max=52*2)

# test for serial correlation in residuals
Box.test(select.resids,lag=8,type='Ljung',fitdf=7)
# test for serial correlation in squared residuals
Box.test((select.resids)^2,lag=8,type='Ljung',fitdf=7)
```


```{r}
nfore = length(temp.test)
forecast_arima_rolling = NULL
forecast_se_arima_rolling = NULL
for(f in 1: nfore){
    ## Fit models
    data = temp.train
    if(f>2)
       data = c(temp.train,temp.test[1:(f-1)])
    model_arima = arima(data, order=c(4,0,3))
    outpred_daily = as.vector(predict(model_arima,n.ahead=1))
    forecast_arima_rolling = c(forecast_arima_rolling, outpred_daily$pred)
    forecast_se_arima_rolling = c(forecast_se_arima_rolling, outpred_daily$se)
}

timevol = time(temp)
ubound = forecast_arima_rolling+1.96*forecast_se_arima_rolling
lbound = forecast_arima_rolling-1.96*forecast_se_arima_rolling
ymin = min(lbound)
ymax = max(ubound)

n = length(temp)
par(mfrow=c(1,1))
plot(timevol[(n-100):n],temp[(n-100):n],type="l", ylim=c(0,40), xlab="Time", ylab="Temperature", main = "ARIMA Weekly Data")
points(timevol[(nfit+1):n],forecast_arima_rolling,col="red")
lines(timevol[(nfit+1):n],ubound,lty=3,lwd= 2, col="blue")
lines(timevol[(nfit+1):n],lbound,lty=3,lwd= 2, col="blue")
```
```{r}
# nfore = length(temp.test)
# forecast_arima_rolling_2 = NULL
# forecast_se_arima_rolling_2 = NULL
# for(f in seq(1,nfore,2)){
#     ## Fit models
#     data = temp.train
#     if(f>2)
#        data = c(temp.train,temp.test[1:(f-1)])
#     model_arima_2 = arima(data, order=c(4,0,3))
#     outpred_daily_2 = as.vector(predict(model_arima_2,n.ahead=2))
#     forecast_arima_rolling_2 = c(forecast_arima_rolling_2, outpred_daily_2$pred)
#     forecast_se_arima_rolling_2 = c(forecast_se_arima_rolling_2, outpred_daily_2$se)
# }

timevol = time(temp)
ubound = forecast_arima_rolling_2+1.96*forecast_se_arima_rolling_2
lbound = forecast_arima_rolling_2-1.96*forecast_se_arima_rolling_2
ymin = min(lbound)
ymax = max(ubound)

n = length(temp)
par(mfrow=c(1,1))
plot(timevol[(n-100):n],temp[(n-100):n],type="l", ylim=c(0,40), xlab="Time", ylab="Temperature", main = "ARIMA Weekly Data (2-week rolling)")
points(timevol[(nfit+1):n],forecast_arima_rolling_2,col="red")
lines(timevol[(nfit+1):n],ubound,lty=3,lwd= 2, col="blue")
lines(timevol[(nfit+1):n],lbound,lty=3,lwd= 2, col="blue")
```

```{r}
n = length(temp)
n_forward=28
nfit = n-n_forward

forecast_arima = as.vector(predict(daily_model,n.ahead=28))
timevol = time(temp)
ubound = forecast_arima$pred+1.96*forecast_arima$se
lbound = forecast_arima$pred-1.96*forecast_arima$se
ymin = min(lbound)
ymax = max(ubound)

n = length(temp)
par(mfrow=c(1,1))
plot(timevol[(n-100):n],temp[(n-100):n],type="l", ylim=c(0,40), xlab="Time", ylab="Temperature", main = "ARIMA Weekly Data (no rolling)")
points(timevol[(nfit+1):n],forecast_arima$pred,col="red")
lines(timevol[(nfit+1):n],ubound,lty=3,lwd= 2, col="blue")
lines(timevol[(nfit+1):n],lbound,lty=3,lwd= 2, col="blue")
```



```{r}
#SARIMA
norder=10
sorder=2
p = c(1:norder)-1; q = c(1:norder)-1
sp = c(1:norder)-1; sq = c(1:norder)-1

aic_sarima_11=matrix(0,norder,norder)
for(i in 1:norder){
  for(j in 1:norder){
    #cat(i,j,'\n')
    tryCatch({
      sarima_model_select3 = astsa::sarima(temp.train, p[i], 0, q[j], 1, 0, 1, 52, details=FALSE, Model=FALSE)
                aic_sarima_11[i,j] = sarima_model_select3$AIC},
                error=function(e) {}
            )
  
  }  
}

aic_sarima_11[which(aic_sarima_11==0)] = NA
which(aic_sarima_11 == min(aic_sarima_11,na.rm=T), arr.ind = TRUE)
aic_sarima_11[which(aic_sarima_11 == min(aic_sarima_11,na.rm=T), arr.ind = TRUE)]

aic_sarima_11_diff=matrix(0,norder,norder)
for(i in 1:norder){
  for(j in 1:norder){
    #cat(i,j,'\n')
    tryCatch({
      sarima_model_select3 = astsa::sarima(temp.train, p[i], 1, q[j], 1, 0, 1, 52, details=FALSE, Model=FALSE)
                aic_sarima_11_diff[i,j] = sarima_model_select3$AIC},
                error=function(e) {}
            )
  
  }  
}

aic_sarima_11_diff[which(aic_sarima_11_diff==0)] = NA
which(aic_sarima_11_diff == min(aic_sarima_11_diff,na.rm=T), arr.ind = TRUE)
aic_sarima_11_diff[which(aic_sarima_11_diff == min(aic_sarima_11_diff,na.rm=T), arr.ind = TRUE)]
```



```{r}
sarima.temp = astsa::sarima(temp.train, 6,0,10, 1,0,1, 52)
sarima_for = sarima.for(temp.train, 28, 6,0,10, 1,0,1, 52)
pred_SARIMA = sarima_for$pred
```


```{r}
n = length(temp)
n_forward=28
nfit = n-n_forward

timevol = time(temp)
ubound = sarima_for$pred+1.96*sarima_for$se
lbound = sarima_for$pred-1.96*sarima_for$se
ymin = min(lbound)
ymax = max(ubound)

n = length(temp)
par(mfrow=c(1,1))
plot(timevol[(n-100):n],temp[(n-100):n],type="l", ylim=c(0,40), xlab="Time", ylab="Temperature", main = "SARIMA Weekly Data (no rolling)")
points(timevol[(nfit+1):n],sarima_for$pred,col="red")
lines(timevol[(nfit+1):n],ubound,lty=3,lwd= 2, col="blue")
lines(timevol[(nfit+1):n],lbound,lty=3,lwd= 2, col="blue")

sarima_resid <- residuals(sarima.temp$fit)

acf(sarima_resid,main="Residuals - Weekly Data",na.action = na.pass,lag.max=52*2)
acf(sarima_resid^2,main="SARIMA Squared Residuals - Weekly Data",na.action = na.pass,lag.max=52*2)

# test for serial correlation in residuals
Box.test(sarima_resid,lag=19,type='Ljung',fitdf=18)
# test for serial correlation in squared residuals
Box.test((sarima_resid)^2,lag=19,type='Ljung',fitdf=18)
```


```{r}
#SARIMA Rolling prediction

nfore = length(temp.test)
forecast_sarima_rolling_2 = NULL
forecast_se_sarima_rolling_2 = NULL
for(f in seq(1,nfore,2)){
    ## Fit models
    data = temp.train
    if(f>2)
       data = c(temp.train,temp.test[1:(f-1)])
    #model_arima_2 = arima(data, order=c(4,0,3))
    #outpred_daily_2 = as.vector(predict(model_arima_2,n.ahead=2))
    outpred_daily_3 = sarima.for(data, 2, 6,0,10, 1,0,1, 52)
    forecast_sarima_rolling_2 = c(forecast_sarima_rolling_2, outpred_daily_3$pred)
    forecast_se_sarima_rolling_2 = c(forecast_se_sarima_rolling_2, outpred_daily_3$se)
}

timevol = time(temp)
ubound = forecast_sarima_rolling_2+1.96*forecast_se_sarima_rolling_2
lbound = forecast_sarima_rolling_2-1.96*forecast_se_sarima_rolling_2
ymin = min(lbound)
ymax = max(ubound)

n = length(temp)
par(mfrow=c(1,1))
plot(timevol[(n-100):(n-2)],temp[(n-100):(n-2)],type="l", ylim=c(0,40), xlab="Time", ylab="Temperature", main = "SARIMA Weekly Data (2-week rolling)")
points(timevol[(nfit+1):(n-2)],forecast_sarima_rolling_2,col="red")
lines(timevol[(nfit+1):(n-2)],ubound,lty=3,lwd= 2, col="blue")
lines(timevol[(nfit+1):(n-2)],lbound,lty=3,lwd= 2, col="blue")

```


```{r}
#SARIMA
norder=10
sorder=2
p = c(1:norder)-1; q = c(1:norder)-1
sp = c(1:norder)-1; sq = c(1:norder)-1
temp.train.log = log(temp.train+(8/3))

aic_sarima_11_log=matrix(0,norder,norder)
for(i in 1:norder){
  for(j in 1:norder){
    #cat(i,j,'\n')
    tryCatch({
      sarima_model_select3 = astsa::sarima(temp.train.log, p[i], 0, q[j], 1, 0, 1, 52, details=FALSE, Model=FALSE)
                aic_sarima_11_log[i,j] = sarima_model_select3$AIC},
                error=function(e) {}
            )
  
  }  
}

aic_sarima_11_log[which(aic_sarima_11_log==0)] = NA
which(aic_sarima_11_log == min(aic_sarima_11_log,na.rm=T), arr.ind = TRUE)
aic_sarima_11_log[which(aic_sarima_11_log == min(aic_sarima_11_log,na.rm=T), arr.ind = TRUE)]

aic_sarima_11_diff_log=matrix(0,norder,norder)
for(i in 1:norder){
  for(j in 1:norder){
    #cat(i,j,'\n')
    tryCatch({
      sarima_model_select3 = astsa::sarima(temp.train.log, p[i], 1, q[j], 1, 0, 1, 52, details=FALSE, Model=FALSE)
                aic_sarima_11_diff_log[i,j] = sarima_model_select3$AIC},
                error=function(e) {}
            )
  
  }  
}

aic_sarima_11_diff_log[which(aic_sarima_11_diff_log==0)] = NA
which(aic_sarima_11_diff_log == min(aic_sarima_11_diff_log,na.rm=T), arr.ind = TRUE)
aic_sarima_11_diff_log[which(aic_sarima_11_diff_log == min(aic_sarima_11_diff_log,na.rm=T), arr.ind = TRUE)]
```



```{r}
sarima.temp.log = astsa::sarima(temp.train.log, 3,0,10, 1,0,1, 52)
sarima_for.log = sarima.for(temp.train.log, 28, 3,0,10, 1,0,1, 52)
pred_SARIMA.log = sarima_for.log$pred

temp.log = log(temp+(8/3))
n = length(temp)
n_forward=28
nfit = n-n_forward

timevol = time(temp)
ubound = sarima_for.log$pred+1.96*sarima_for.log$se
lbound = sarima_for.log$pred-1.96*sarima_for.log$se
ymin = min(lbound)
ymax = max(ubound)

n = length(temp)
par(mfrow=c(1,1))
plot(timevol[(n-100):n],temp.log[(n-100):n],type="l", ylim=c(ymin,ymax), xlab="Time", ylab="Temperature", main = "Log SARIMA Weekly Data (no-rolling)")
points(timevol[(nfit+1):n],sarima_for.log$pred,col="red")
lines(timevol[(nfit+1):n],ubound,lty=3,lwd= 2, col="blue")
lines(timevol[(nfit+1):n],lbound,lty=3,lwd= 2, col="blue")

sarima_resid.log <- residuals(sarima.temp.log$fit)

acf(sarima_resid.log,main="Residuals - Weekly Data",na.action = na.pass,lag.max=52*4)
acf(sarima_resid.log^2,main="Log SARIMA Squared Residuals - Weekly Data",na.action = na.pass,lag.max=52*4)

# test for serial correlation in residuals
Box.test(sarima_resid,lag=16,type='Ljung',fitdf=15)
# test for serial correlation in squared residuals
Box.test((sarima_resid)^2,lag=16,type='Ljung',fitdf=15)
```


```{r}
#SARIMA Rolling prediction

nfore = length(temp.test)
forecast_sarima_rolling_2.log = NULL
forecast_se_sarima_rolling_2.log = NULL
for(f in seq(1,nfore,2)){
    ## Fit models
    data = temp.train.log
    if(f>2)
       data = c(temp.train.log,log(temp.test[1:(f-1)]+10))
    #model_arima_2 = arima(data, order=c(4,0,3))
    #outpred_daily_2 = as.vector(predict(model_arima_2,n.ahead=2))
    outpred_daily_4 = sarima.for(data, 2, 3,0,10, 1,0,1, 52)
    forecast_sarima_rolling_2.log = c(forecast_sarima_rolling_2.log, outpred_daily_4$pred)
    forecast_se_sarima_rolling_2.log = c(forecast_se_sarima_rolling_2.log, outpred_daily_4$se)
}

timevol = time(temp)
ubound = forecast_sarima_rolling_2.log+1.96*forecast_se_sarima_rolling_2.log
lbound = forecast_sarima_rolling_2.log-1.96*forecast_se_sarima_rolling_2.log
ymin = min(temp.log[(n-100):n])
ymax = max(ubound)

n = length(temp)
par(mfrow=c(1,1))
plot(timevol[(n-100):n],temp.log[(n-100):n],type="l", ylim=c(ymin,ymax), xlab="Time", ylab="Temperature", main = "Log SARIMA Weekly Data (2-week rolling)")
points(timevol[(nfit+1):n],forecast_sarima_rolling_2.log,col="red")
lines(timevol[(nfit+1):n],ubound,lty=3,lwd= 2, col="blue")
lines(timevol[(nfit+1):n],lbound,lty=3,lwd= 2, col="blue")
```

```{r}
# Prediction Comparison

mape <- function(y, y_pred) {
  mape <- mean(abs((y-y_pred)/y))
  return(mape)
}

pm <- function(obs, pred) {
  pm <- sum((pred-obs)^2)/sum((obs-mean(obs))^2)
  return(pm)
}

#MAPE
m1 = mape(temp.test,forecast_arima_rolling)
m2 = mape(temp.test,forecast_arima_rolling_2)
m3 = mape(temp.test,forecast_arima$pred)
m4 = mape(temp.test,sarima_for$pred)
m5 = mape(temp.test[1:26],forecast_sarima_rolling_2)
m6 = mape(log(temp.test+(8/3)),sarima_for.log$pred)
m7 = mape(log(temp.test+(8/3)),forecast_sarima_rolling_2.log)

#PM
p1 = pm(temp.test,forecast_arima_rolling)
p2 = pm(temp.test,forecast_arima_rolling_2)
p3 = pm(temp.test,forecast_arima$pred)
p4 = pm(temp.test,sarima_for$pred)
p5 = pm(temp.test[1:26],forecast_sarima_rolling_2)
p6 = pm(log(temp.test+(8/3)),sarima_for.log$pred)
p7 = pm(log(temp.test+(8/3)),forecast_sarima_rolling_2.log)
```


```{r}
prediction_mat = NULL
for (i in 2:7){
  prediction_mat = rbind(prediction_mat,data.frame(eval(as.symbol(paste("m",i,sep=''))),eval(as.symbol(paste("p",i,sep='')))))
}
rownames(prediction_mat) <- c("ARIMA (2-week rolling)","ARIMA (non-rolling)","SARIMA (non-rolling)","SARIMA (2-week rolling)","SARIMA log transformed (no-rolling)","SARIMA log transformed (2-week rolling)")
colnames(prediction_mat) <- c("MAPE", "PM")
prediction_mat
```


```{r}
ts.sim <- arima.sim(list(order = c(4,0,3), ar = c(1.2990,0.6765,-1.3089,0.3135), ma = c(-0.9686,-0.9860,0.9827)), n = 200)
ts.plot(ts.sim)
```

